---
title: "Advance R Programming"
author: "SUYOG TIMALSINA"
output: html_document
date: "2025-02-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Built-in R Feature and Functions
```{r }
# seq() : Create sequence
seqA <- seq(0,10, by = 2)
seqA

#sort() : sort a vector
v <- c(1111,22,3,4,5,6,998,98,55)
sort(v)

sort(v, decreasing = TRUE)

# rev() : Reverse elements in object
rev(v)

#as.: Converting the data type
as.list(v)

as.matrix(v)
```

# 2. Apply
* `sapply` is used to do list and function. The result is always a list.
* `lapply` works on both vectors and lists.It simplifies the result. If the result of applying the function is a single value for each element, it will return a vector instead of a list.
```{r}
v <- c(1,2,3,4,5)
addrand <- function(x) {
  ran <- sample(1:100,1)
  return(x + ran)
}

result <- lapply(v,addrand)
result

v1 <- 1:5

times2 <- function(num) {
  return(num * 2)
}
print(v)
result <- sapply(v1, times2)
result
```

*Anonymous Functions
```{r}
v <- 1:5
result <- sapply(v, function(num){num * 2})
result
```

# 2. Data Manipulation
```{r}

library("dplyr")


library(nycflights13)

head(flights)
```

```{r results='hide'}
summary(flights)

```

* `filter`
```{r}
filter(flights,month == "11", day == 3, carrier == "AA")
```

* `slice`
* helps to select the rows
```{r}
slice(flights,1:10)
```

* `arrange`
```{r}
head(arrange(flights,year,month,day,arr_time))
```

* `select`
```{r}
head(select(flights,carrier,arr_time,month))
```

* `rename`
```{r}
head(rename(flights,airline_carrier = carrier))
```

* `distinct`
*  removes those duplicates, leaving only the unique carriers
```{r}
distinct(select(flights,carrier))
```

* `mutate`
to new  column
```{r}
head(mutate(flights, new_col = arr_delay- dep_delay ))
```

* transmute 
just give the output of the new column
```{r}
head(transmute(flights, new_col = arr_delay- dep_delay ))
```

* `summarize`
```{r}
summarize(flights,avg_air_time = mean(air_time,na.rm = TRUE))

summarize(flights,total_air_time = sum(air_time,na.rm = TRUE))
```

```{r}

```

* `sample_n`
```{r}
sample_n(flights,10)
```

* `sample_frac`
```{r}
sample_frac(flights,0.1)
```

# 3. Pipe Operator
* useful to use multiple function 
```{r}
library(dplyr)
df <- mtcars
head(df)

#Nesting
# result <- arrange(sample_n(filter(df, mpg > 20), size = 5))

# Pipe Operaztror
# data %>% op1 %>%
result <- df %>% filter(mpg > 20) %>% sample_n(size = 5) %>% arrange(desc(mpg))
result


```

# 4. tidyr package
* `tidyr package` and `data.table` install
```{r}
library(tidyr)
library(data.table)

# Example of data set
comp <- c(1,1,1,2,2,2,3,3,3)
yr <- c(1998,1999,2000,1998,1999,2000,1998,1999,2000)
q1 <- runif(9, min=0, max=100)
q2 <- runif(9, min=0, max=100)
q3 <- runif(9, min=0, max=100)
q4 <- runif(9, min=0, max=100)

df <- data.frame(comp=comp,year=yr,Qtr1 = q1,Qtr2 = q2,Qtr3 = q3,Qtr4 = q4)
df
```
* gather() 
The gather() function will collapse multiple columns into key-pair values. The data frame above is considered wide since the time variable (represented as quarters) is structured such that each quarter represents a variable. To re-structure the time component as an individual variable, we can gather each quarter within one column variable and also gather the values associated with each quarter in a second column variable.
```{r}
# using pipe operator
head((df %>% gather(Quarter , Revene, Qtr1:Qtr4)))

```

* spread()
This is the complement of gather(), which is why its called spread():
```{r}
library(tidyr)
library(dplyr)

# Create a sample dataset in long format
stocks_long <- data.frame(
  time = rep(as.Date('2009-01-01') + 0:9, each = 3),  # Repeat each date 3 times (for X, Y, Z)
  stock_type = rep(c("X", "Y", "Z"), times = 10),     # Stock type (X, Y, Z)
  value = c(rnorm(10, 0, 1), rnorm(10, 0, 2), rnorm(10, 0, 4))  # Random values
)

# Print long format
head(stocks_long)

# Using long to wide using spread()
stocks_wide <- spread(stocks_long, key = stock_type, value = value)
print(stocks_wide)



```

* unite() 
Unite is a convenience function to paste together multiple columns into one.
```{r}
library(tidyr)

data <- data.frame(
  first_name = c("John", "Jane", "David"),
  last_name = c("Doe", "Smith", "Brown")
)

print(data)

# Using unite() 
data_united <- unite(data, full_name, first_name, last_name, sep = " ")
print(data_united)


```









